---
title: "Crashes Project 10_12"
author: "Paige Singla"
date: "2023-10-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(sf)
library(dplyr)
library(tidygeocoder)
library(tidyverse)
```

```{r}
traffic <- read.csv("Average_Daily_Traffic_Counts.csv")
crashes <- read.csv("Traffic_Crashes.csv")
speed <- read.csv("Speed_Camera_Violations.csv")
red <- read.csv("Red_Light_Camera_Violations.csv")
```

```{r}
file <- "https://data.cityofchicago.org/resource/qqq8-j68g.geojson"
chicago <- read_sf(file)
class(chicago)
st_crs(chicago)
plot(chicago)


grid <- chicago %>% 
  st_make_grid(cellsize = c(.01,0.01), square = TRUE) %>% # grid of points
  st_intersection(chicago) 

plot(grid) #length 766
```
## Traffic

```{r}
df_traffic_sf <- st_as_sf(traffic,
                  coords = c('Longitude','Latitude'), crs = st_crs(4326))


within_traffic <- st_within(df_traffic_sf,grid)
for (i in 1:length(within_traffic)){
if (length(within_traffic[[i]]) == 0){within_traffic[[i]] <- -99}
}


(within_traffic)
num_traffic <- (table(unlist(within_traffic)))
(str(within_traffic))
table(unlist(lapply(within_traffic, length)))


```

```{r}
traffic2 <- traffic
traffic2$grid_num <- unlist(within_traffic) #map every element to a cell number and group by cell number
#length(unlist(within_traffic))
grid_num <- seq(1,length(grid))

df_end <- data.frame(grid_num) #create df with variable grid 1-1596


traffic3 <- traffic2 %>% group_by(grid_num) %>% summarise(vehichle_volume = sum(Total.Passing.Vehicle.Volume))
df_end <- merge(df_end, traffic3, by.x = "grid_num", by.y = "grid_num", all.x = TRUE)
df_end
#mean(is.na(df_end))
```


## red
```{r}
red2 <- red %>% filter(is.na(LATITUDE))
head(red2)
```
```{r}
red2 <- distinct(red2, CAMERA.ID, .keep_all = TRUE)
lat_longs_red <- red2 %>%
  mutate(addy = paste0(ADDRESS, ", Chicago, IL")) %>% 
  geocode(addy, method = 'arcgis', lat = LATITUDE , long = LONGITUDE)

lat_longs_red <- lat_longs_red[-20,] # this row had no camera id and one violation, it has an address that is associated with id 2054
lat_longs_red_new <- subset(lat_longs_red, select = c('CAMERA.ID','LATITUDE...12','LONGITUDE...13'))#subset data to have just id lat long

unique_red <- distinct(red, CAMERA.ID, .keep_all = TRUE)
head(unique_red)

unique_red <- merge(unique_red , lat_longs_red_new, by = 'CAMERA.ID', all.x = TRUE) # adding on the new latitude
unique_red$LATITUDE[is.na(unique_red$LATITUDE)] <- unique_red$LATITUDE...12[is.na(unique_red$LATITUDE)] # merging latitude

unique_red$LONGITUDE[is.na(unique_red$LONGITUDE)] <- unique_red$LONGITUDE...13[is.na(unique_red$LONGITUDE)] # merging longitude
unique_red <- unique_red[-366,] # this row had no camera id and one violation, it has an address that is associated with id 2054
unique_red <- unique_red[-359,]
```

```{r}

df_red = na.omit(unique_red[1:365,c('LATITUDE','LONGITUDE')])
df_red_sf <- st_as_sf(df_red,
                  coords = c('LONGITUDE','LATITUDE'), crs = st_crs(4326))


within_red <- st_within(df_red_sf,grid)

(within_red)
num_red <- (table(unlist(within_red)))
(str(within_red))
```

map every element to a cell number and group by cell number
create df with variable grid 1-1596 then merge within_red groupby grid for unique
og file merge camera id group by grid number
map from camera id to grid number left join in to our merged tata and merge by camera id, take merged data set and group by grid number but sum by violations

```{r}
unique_red$grid_num <- unlist(within_red) #map every element to a cell number and group by cell number



red3 <- unique_red %>% group_by(grid_num) %>% summarise(num_red = n())
#red4 <- red %>% group_by()
df_end <- merge(df_end, red3, by.x = "grid_num", by.y = "grid_num", all.x = TRUE)
df_end


```


*Red Total Violations*
```{r}
red_A <- red
red_A <- merge(red_A , lat_longs_red_new, by = 'CAMERA.ID', all.x = TRUE) # adding on the new latitude
red_A$LATITUDE[is.na(red_A$LATITUDE)] <- red_A$LATITUDE...12[is.na(red_A$LATITUDE)] # merging latitude

red_A$LONGITUDE[is.na(red_A$LONGITUDE)] <- red_A$LONGITUDE...13[is.na(red_A$LONGITUDE)] # merging longitude
```

```{r}
df_red_A = na.omit(red_A[1:878585,c('LATITUDE','LONGITUDE')])
df_red_A_sf <- st_as_sf(df_red_A,
                  coords = c('LONGITUDE','LATITUDE'), crs = st_crs(4326))


within_red_A <- st_within(df_red_A_sf,grid)

for (i in 1:length(within_red_A)){
if (length(within_red_A[[i]]) == 0){within_red_A[[i]] <- -99}
}

(within_red_A)
num_red_A <- (table(unlist(within_red_A)))
(str(within_red_A))
```


```{r}
red_A <- red_A[ -c(6,7,11:12) ]
red_A <- red_A %>% drop_na(LONGITUDE)
red_A$grid_num <- unlist(within_red_A) #map every element to a cell number and group by cell number



red_B <- red_A %>% group_by(grid_num) %>% summarise(red_vio = sum(VIOLATIONS))
#red4 <- red %>% group_by()
df_end <- merge(df_end, red_B, by.x = "grid_num", by.y = "grid_num", all.x = TRUE)
df_end
```
PAIGE ASK ABOUT THIS ERROR!!!!!!!!




## SPEED
```{r}
speed2 <- speed %>% filter(is.na(LATITUDE))
head(speed2)
```

```{r}
speed2 <- distinct(speed2, CAMERA.ID, .keep_all = TRUE)
lat_longs_speed <- speed2 %>%
  mutate(addy = paste0(ADDRESS, ", Chicago, IL")) %>% 
  geocode(addy, method = 'arcgis', lat = LATITUDE , long = LONGITUDE)

lat_longs_speed_new <- subset(lat_longs_speed, select = c('CAMERA.ID','LATITUDE...11','LONGITUDE...12'))#subset data to have just id lat long

unique_speed <- distinct(speed, CAMERA.ID, .keep_all = TRUE)
head(unique_speed)

unique_speed <- merge(unique_speed , lat_longs_speed_new, by = 'CAMERA.ID', all.x = TRUE) # adding on the new latitude and longitude
unique_speed$LATITUDE[is.na(unique_speed$LATITUDE)] <- unique_speed$LATITUDE...11[is.na(unique_speed$LATITUDE)] # merging latitude
unique_speed$LONGITUDE[is.na(unique_speed$LONGITUDE)] <- unique_speed$LONGITUDE...12[is.na(unique_speed$LONGITUDE)] # merging longitude

```

```{r}
df_speed = na.omit(unique_speed[1:365,c('LATITUDE','LONGITUDE')])
df_speed_sf <- st_as_sf(df_speed,
                  coords = c('LONGITUDE','LATITUDE'), crs = st_crs(4326))


within_speed <- st_within(df_speed_sf,grid)

(within_speed)
num_speed <- (table(unlist(within_speed)))
(str(within_speed))
```

map every element to a cell number and group by cell number
create df with variable grid 1-1596 then merge within_speed groupby grid for unique
og file merge camera id group by grid number
map from camera id to grid number left join in to our merged tata and merge by camera id, take merged data set and group by grid number but sum by violations

```{r}
unique_speed$grid_num <- within_speed #map every element to a cell number and group by cell number

speed3 <- unique_speed %>% group_by(grid_num) %>% summarise(num_speed = n())
df_end <- merge(df_end, speed3, by.x = "grid_num", by.y = "grid_num", all.x = TRUE)
df_end
```
*Speed Total Violations*
```{r}
speed_A <- speed
speed_A <- merge(speed_A , lat_longs_red_new, by = 'CAMERA.ID', all.x = TRUE) # adding on the new latitude
speed_A$LATITUDE[is.na(speed_A$LATITUDE)] <- speed_A$LATITUDE...12[is.na(speed_A$LATITUDE)] # merging latitude

speed_A$LONGITUDE[is.na(speed_A$LONGITUDE)] <- speed_A$LONGITUDE...13[is.na(speed_A$LONGITUDE)] # merging longitude
```

```{r}
df_speed_A = na.omit(speed_A[1:878585,c('LATITUDE','LONGITUDE')])
df_speed_A_sf <- st_as_sf(df_speed_A,
                  coords = c('LONGITUDE','LATITUDE'), crs = st_crs(4326))


within_speed_A <- st_within(df_speed_A_sf,grid)

(within_speed_A) 
num_speed_A <- (table(unlist(within_speed_A)))
(str(within_speed_A))
```

```{r}
speed_A <- speed_A[ -c(5,6,10:11) ]
speed_A <- speed_A %>% drop_na(LONGITUDE)
speed_A$grid_num <- unlist(within_speed_A) #map every element to a cell number and group by cell number



speed_B <- speed_A %>% group_by(grid_num) %>% summarise(speed_vio = sum(VIOLATIONS))
#red4 <- red %>% group_by()
df_end <- merge(df_end, speed_B, by.x = "grid_num", by.y = "grid_num", all.x = TRUE)
df_end
```


##CRASHES

```{r}
crashes2 <- (crashes)
```

```{r}
df_crashes = na.omit(crashes2[1:759443,c('LATITUDE','LONGITUDE')])
df_sf <- st_as_sf(df_crashes,
                  coords = c('LONGITUDE','LATITUDE'), crs = st_crs(4326))


within_crashes <- st_within(df_sf,grid)


```

map every element to a cell number and group by cell number
create df with variable grid 1-1596 then merge within_crashes groupby grid for unique
og file merge camera id group by grid number
map from camera id to grid number left join in to our merged tata and merge by camera id, take merged data set and group by grid number but sum by violations

```{r}
df_crashes$grid_num <- within_crashes #map every element to a cell number and group by cell number

crashes3 <- df_crashes %>% group_by(grid_num) %>% summarise(num_crashes = n())
df_end <- merge(df_end, crashes3, by.x = "grid_num", by.y = "grid_num", all.x = TRUE)
df_end <- df_end %>% replace(is.na(.),0)

df_end

# filtering by year 2019
df_crashes$CRASH_DATE <- as.Date(df_crashes$CRASH_DATE, format =  "%m/%d/%Y %H:%M:%S")
df_crashes$YEAR <- as.numeric((format(df_crashes$CRASH_DATE, "%Y")))
crashes_2019 <- df_crashes %>% filter(YEAR == 2019)
crashes_2019 <- df_crashes %>% group_by(grid_num) %>% summarise(num_crashes = n())
df_2019 <- merge(df_2019, crashes_2019, by.x = "grid_num", by.y = "grid_num", all.x = TRUE)
#df_2019 <- df_2019 %>% replace(is.na(.),0)

df_2019
```

pick some cut off 
build a model num crashes per day as a reponse of daily thingd year 2019

do code exactly the same output is a record and cell, add cell number to dataset instead of counting with n = n() now we would do tot vio = sum(violations)

average traffic or max, max

probably poisson distribution in glm iid assumption for now

link functions - look this up

correlation between grids

if camerea id 

filtering by year 2019
```{r}
df_2019 <- data.frame(grid_num) 
df_2019 <- merge(df_2019, traffic3, by.x = "grid_num", by.y = "grid_num", all.x = TRUE)
```

```{r}
#df %>% filter(between(date_column, as.Date('2022-01-20'), as.Date('2022-02-20')))
crashes_3 <- (crashes)
crashes_3$CRASH_DATE <- as.Date(crashes_3$CRASH_DATE, format =  "%m/%d/%Y %H:%M:%S")
crashes_3$CRASH_DATE <- as.numeric((format(crashes_3$CRASH_DATE, "%Y")))
crashes_2019 <- crashes_3 %>% filter(CRASH_DATE == 2019)

df_crashes_2019 = na.omit(crashes_2019[1:117762,c('LATITUDE','LONGITUDE')])
df_sf_2019 <- st_as_sf(df_crashes_2019,
                  coords = c('LONGITUDE','LATITUDE'), crs = st_crs(4326))


within_crashes_2019 <- st_within(df_sf_2019,grid)
```

```{r}
df_crashes_2019$grid_num <- within_crashes_2019 #map every element to a cell number and group by cell number

df_crashes_2019_new <- df_crashes_2019 %>% group_by(grid_num) %>% summarise(num_crashes_2019 = n())
df_2019 <- merge(df_2019, df_crashes_2019_new, by.x = "grid_num", by.y = "grid_num", all.x = TRUE)

df_2019
```
*Red Total Violations 2019*

```{r}
red_B <- red
red_B$VIOLATION.DATE <- as.Date(red_B$VIOLATION.DATE, format =  "%m/%d/%Y")
red_B <- red_B %>% filter(between(VIOLATION.DATE, as.Date('2019-01-01'), as.Date('2019-12-31')))
red_B_2019 <- red_B %>% group_by(CAMERA.ID) %>% summarise(num_days = n())

red_B <- merge(red_B , red_B_2019, by = 'CAMERA.ID', all.x = TRUE)

red_B_new <- subset(red_B, num_days > 300)
```


```{r}
red_B_new <- merge(red_B_new , lat_longs_red_new, by = 'CAMERA.ID', all.x = TRUE) # adding on the new latitude
red_B_new$LATITUDE[is.na(red_B_new$LATITUDE)] <- red_B_new$LATITUDE...12[is.na(red_B_new$LATITUDE)] # merging latitude

red_B_new$LONGITUDE[is.na(red_B_new$LONGITUDE)] <- red_B_new$LONGITUDE...13[is.na(red_B_new$LONGITUDE)] # merging longitude
```

```{r}
df_red_B_new = na.omit(red_B_new[1:71334,c('LATITUDE','LONGITUDE')])
df_red_B_new_sf <- st_as_sf(df_red_B_new,
                  coords = c('LONGITUDE','LATITUDE'), crs = st_crs(4326))


within_red_B_new <- st_within(df_red_B_new_sf,grid)

for (i in 1:length(within_red_B_new)){
if (length(within_red_B_new[[i]]) == 0){within_red_B_new[[i]] <- -99}
}

(within_red_B_new)
num_red_B_new <- (table(unlist(within_red_B_new)))
(str(within_red_B_new))
```

```{r}
red_B_new <- red_B_new[ -c(12:15) ]
red_B_new <- red_B_new %>% drop_na(LONGITUDE)
red_B_new$grid_num <- unlist(within_red_B_new) #map every element to a cell number and group by cell number



red_C_new <- red_B_new %>% group_by(grid_num) %>% summarise(red_vio_2019 = sum(VIOLATIONS))
#red4 <- red %>% group_by()
df_2019 <- merge(df_2019, red_C_new, by.x = "grid_num", by.y = "grid_num", all.x = TRUE)
df_2019
```


*Speed Total Violations 2019*
```{r}
speed_B <- speed
speed_B$VIOLATION.DATE <- as.Date(speed_B$VIOLATION.DATE, format =  "%m/%d/%Y")
speed_B <- speed_B %>% filter(between(VIOLATION.DATE, as.Date('2019-01-01'), as.Date('2019-12-31')))
speed_B_2019 <- speed_B %>% group_by(CAMERA.ID) %>% summarise(num_days = n())

speed_B <- merge(speed_B , speed_B_2019, by = 'CAMERA.ID', all.x = TRUE)

speed_B_new <- subset(speed_B, num_days > 300)
```



```{r}
df_speed_B_new = na.omit(speed_B_new[1:26924,c('LATITUDE','LONGITUDE')])
df_speed_B_new_sf <- st_as_sf(df_speed_B_new,
                  coords = c('LONGITUDE','LATITUDE'), crs = st_crs(4326))


within_speed_B_new <- st_within(df_speed_B_new_sf,grid)

for (i in 1:length(within_speed_B_new)){
if (length(within_speed_B_new[[i]]) == 0){within_speed_B_new[[i]] <- -99}
}

(within_speed_B_new)
num_speed_B_new <- (table(unlist(within_speed_B_new)))
(str(within_speed_B_new))
```

```{r}
speed_B_new <- speed_B_new[ -c(12:15) ]
speed_B_new <- speed_B_new %>% drop_na(LONGITUDE)
speed_B_new$grid_num <- unlist(within_speed_B_new) #map every element to a cell number and group by cell number



speed_C_new <- speed_B_new %>% group_by(grid_num) %>% summarise(speed_vio_2019 = sum(VIOLATIONS))
#speed4 <- speed %>% group_by()
df_2019 <- merge(df_2019, speed_C_new, by.x = "grid_num", by.y = "grid_num", all.x = TRUE)
df_2019 <- df_2019 %>% replace(is.na(.),0)
df_2019
```


*MODEL*
```{r}
poisson_mod <- glm(num_crashes_2019 ~ red_vio_2019 + speed_vio_2019 + vehichle_volume, family = "poisson", data = df_2019)
summary(poisson_mod)
```

