---
title: "Homework 7"
author: "Jacob Plaza"
date: "2023-04-18"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Apply the McNemar test to Table 8.3 on smoking and birthweight. Interpret.

```{r}


smoke = matrix(c(159,8,22,14), nrow =2)
mcnemar.test(smoke, correct = FALSE)
```

X^2 value is 6.53, and the p-value is equal to 0.01. So we would reject the null hypothesis. Therefore, low birth rate cases are more likely to have mothers who smoke than normal weight controls. 

### 8.3 Explain the diﬀerence between B for marginal model (8.2) and B for subject-speciﬁc model (8.3). Illustrate with Table 8.1 on helping the environment.


For a marginal model, B is the log odds ratio (log(p/1-p)) comparing marginal distributions. For a subject specific model, B is conditional on the subject, which means that the odds of success for yi1 are exp(B) multiplied by the odds of success for yi2. 

### 8.4 Estimate exp(B) and explain why the estimate is valid even though the study observed diabetes, conditional on MI status.

```{r}
oddsratio = (9/37)/(16/82)
oddsratio
```

B would be the log odds ratio, so after exponentiating, the parameter estimate would just be 1.246622.


### 8.6 Explain the following analogy: The McNemar test is to binary response variables as the paired-diﬀerence t test is to normally distributed response variables.

A McNemar test is used when you have matched-pairs data with a binary response. You use it when you want to compare two dependent proportions. A paired difference t-test is a little different, since it examines the mean difference between a pair of measurements. 
